---
title: "Stat 214 project"
author: "Emmanuel Oluloto"
date: "April 19, 2018"
output: html_document
---

<link rel="stylesheet" href="https://www.w3schools.com/w3css/4/w3.css"> 

```{r setup, include=FALSE}
library(lattice)      # For xy plot
library(mosaic)

library(grid)    # For plot grids
library(gridExtra)
library(ggplot2)

library(readr)
library(taRifx)
library(agricolae)
library(lsmeans)

Video_Games_Sales_as_at_22_Dec_2016 <- read_csv("~/R Prog/datasets/Video_Games_Sales_as_at_22_Dec_2016.csv")
Original_Dataset = Video_Games_Sales_as_at_22_Dec_2016 

Original_Dataset$Year_of_Release =  suppressWarnings(as.numeric(Original_Dataset$Year_of_Release)) 
Original_Dataset$User_Score = suppressWarnings(as.numeric(Original_Dataset$User_Score))


# No NAs at all
GameSales_Clean.df = na.omit(Video_Games_Sales_as_at_22_Dec_2016)
colnames(GameSales_Clean.df)

```

## DATASET

I'm working with a table that contains stats for video game sales obtained from Kaggle : https://www.kaggle.com/rush4ratio/video-game-sales-with-ratings

```{r c-list}
colnames(GameSales_Clean.df)

str(GameSales_Clean.df)
```




## Research Questions

####  1. Regression Question  (EX - Q, RESP - Q)
   How do game sales in Japan change over time, with respect to user score?  
   
   Explanatory - Year_of_Release, User_Score* (Quant)    
   Response    - JP_Sales (millions)          (Quant)   

MODEL: 
$$Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_1^2 + \beta_4X_2^2 + \epsilon$$
$$Y = \text{Game Sales in Japan}, X_1 = \text{Year of Release}, X_2 = \text{User Score}$$

  I used an almost-complete second-order model (minus the interaction). I had to test a LOT of models to find the one that worked best. I decided not to factor in the User Count, as that made the model more complex while not really being significant.   


```{r regression, echo=TRUE}

analysis_1 <- Original_Dataset[!(Original_Dataset$Year_of_Release=="N/A"),]
analysis_1 <- Original_Dataset[!(is.na(Original_Dataset$User_Score)),]
#str(analysis_1)

# Change year to numeric
analysis_1$Year_of_Release =  as.numeric(analysis_1$Year_of_Release)
attach(analysis_1)

xyplot(JP_Sales ~ Year_of_Release, main="JP Sales vs Year", pch = 18)

# Original Model
JPSales.lm = lm(JP_Sales ~ Year_of_Release + User_Score + I(User_Score^2) + I(Year_of_Release^2)) 
summary(JPSales.lm)


#Assessing Model Conditions
par(mfrow = c(2,2))
hist(JPSales.lm$residuals, main=NULL)

plot(JPSales.lm$fitted.values, JPSales.lm$residuals)
abline(h=0)

qqnorm(JPSales.lm$residuals, main=NULL)
qqline(JPSales.lm$residuals)
par(mfrow = c(1,1))


# Transforming the data
squrtJP = sqrt(JP_Sales)

TransJP.lm = lm(squrtJP ~ Year_of_Release + User_Score + I(User_Score^2) + I(Year_of_Release^2))
summary(TransJP.lm)

xyplot(squrtJP ~ Year_of_Release, main="Sq-Root JP Sales vs Year", pch = 18)



#Assessing Model Conditions Again
par(mfrow = c(2,2))

hist(TransJP.lm$residuals, main="Histogram of Residuals (Transform)")

plot(TransJP.lm$fitted.values, TransJP.lm$residuals, main = "Fitted values vs Residuals (Transform)")
abline(h=0)

qqnorm(TransJP.lm$residuals, main="QQPlot (Transform)")
qqline(TransJP.lm$residuals)
par(mfrow = c(1,1))


detach(analysis_1)
```




   Conclusion: The year of release and User Score of games do have a relationship with the number of sales in Japan. However, I cannot say this with complete certainty, as the model only explains 5.1% of the variability.

    The model assumptions are also violated. My QQPlot with my original data shows more extreme values than we would expect at one of its ends. The Fits/Residuals plot is also clumped around a certain side of the zero line. I tried every transform on each one of the variables separately and together, and I saw that a square root transform of the Sales worked the best, as the histogram showed a slightly more normal distribution, and the Fits vs Residuals are also a bit more varied around zero. The assumptions were still not met though.






####  2. ANOVA Question  (EX - C, RESP - Q)
   How does the user score change with the game genre?
   
   Explanatory - Genre  
   Response    - User_Score 



There are more than 2 groups (about 12 levels), so a two-sampled t-test does not work out for us.
An ANOVA test would be appropriate!

MODEL: 
$$Y = \mu + \alpha_i  + \epsilon$$
$$Y = \text{User Score}, \alpha_i = \text{Genre Group Effect}$$


```{r ANOVA, echo=TRUE}
analysis_2 <- Original_Dataset[!(is.na(Original_Dataset$User_Score)),]

# Use Genre as Factor
analysis_2$Genre <- as.factor(analysis_2$Genre) 
analysis_2$Developer <- as.character(analysis_2$Developer) 
analysis_2$Rating <- as.factor(analysis_2$Rating) 

attach(analysis_2)
levels(Genre)
levels(Rating)

xyplot(User_Score ~ Genre, main="User Scores vs Genre", pch=16)


UserScore.aov = aov(User_Score ~ Genre)
summary(UserScore.aov)


### Check Model Assumptions ###
par(mfrow = c(2,2))
hist(UserScore.aov$residuals)

plot(UserScore.aov$fitted.values, UserScore.aov$residuals, main = "Fitted values vs Residuals")
abline(h=0)

qqnorm(UserScore.aov$residuals, main = "QQPlot Original")
qqline(UserScore.aov$residuals)
par(mfrow = c(1,1))


# Transformation
score_sq = User_Score^2
TransUserScore.aov = aov(score_sq ~ Genre)

par(mfrow = c(2,2))
hist(TransUserScore.aov$residuals)
plot(TransUserScore.aov$fitted.values, TransUserScore.aov$residuals, main = "(Square) Fitted values vs Residuals")
abline(h=0)
qqnorm(TransUserScore.aov$residuals, main = "(Square) QQPlot for Transform")
qqline(TransUserScore.aov$residuals)
par(mfrow = c(1,1))


summary(TransUserScore.aov)

```



We know that the User Scores vary significantly with each group, but how are the groups different?


```{r}
# Which groups are significantly different?
#FisherTest - LSD.test(Response, Explanatory, df-resid, meansq-resid)
fisher <- LSD.test(score_sq, Genre, 7578, 349)
fisher

# Tukey
TukeyHSD(TransUserScore.aov)


detach(analysis_2)
```

Conclusion: User Scores do vary by game genre. The relationship from the Fisher's test can be summarized below.

+---------------+--------------+--------------+------+
|     A         |       B      |       C      | D    |
+===============+==============+==============+======+
| Role-Playing  | Strategy     | Puzzle       | Misc | 
+---------------+--------------+--------------+------+
|               | Platform     | Adventure    |      |
+---------------+--------------+--------------+------+
|               | Fighting     | Simulation   |      |
+---------------+--------------+--------------+------+
|               | Adventure    | Racing       |      |
+---------------+--------------+--------------+------+
|               | Simulation   | Shooter      |      | 
+---------------+--------------+--------------+------+
|               | Puzzle       | Action       |      |
+---------------+--------------+--------------+------+
|               |              | Sports       |      | 
+---------------+--------------+--------------+------+

The model assumptions are not extremely well met, as the histogram does not show a very normal plot, and the QQPlot shows more extreme deviations than we would expect from a normal distribution at its ends. Transforming the data by squaring the User Score only helped a little. 




####  3. Logistic Regression Question  (EX - Q, RESP - C)
   Do critics give games different scores change before and after 2010?
   Is a highly critiqued game more likely to be before or after 2010?
   
   Explanatory - Critic Score  
   Response    - Before and After 2010  

MODEL: 
$$\pi = \frac {e^{\beta_0 + \beta_1X }} {1 + e^{\beta_0 + \beta_1X }}$$
$$\pi = \text{P(1), Probability that game was made after 2010}, X = \text{Critic score}$$

```{r logistic, echo=TRUE}
year = as.factor(analysis_2$Year_of_Release)
levels(year)
table(year)


After2010 = as.numeric(analysis_2$Year_of_Release > 2010) # 0 : <= 2010, 1 : > 2010 
analysis_3 = analysis_2
analysis_3$After2010 = After2010
attach(analysis_3)

YearMade.lm = lm(After2010 ~ Critic_Score)
summary(YearMade.lm)

plot(After2010 ~ Critic_Score, data = analysis_3)
abline(YearMade.lm)


# Using a General Linear Model
YearGen.glm = glm(After2010 ~ Critic_Score, family=binomial, data = analysis_3)
summary(YearGen.glm)

b0 = YearGen.glm$coefficients[1] # intercept
b1 = YearGen.glm$coefficients[2] # slope
plot(analysis_3$Critic_Score, jitter(analysis_3$After2010,amount=.05),ylab="After 2010")
curve(exp(b0+b1*x)/(1+exp(b0+b1*x)),add=T,col="red")
title(main="Data & Predicted Values from Logistic Regression Model")



```


Conclusion: The critic score is a good predictor of the probability that the game was made at a certain time period.
Looking at the plot from the general linear model, we can see that the probability of a game being made after 2010 increases with a better critic score.

<style>

hr {
	border-top: 4px double #8c8b8b;
	text-align: center;
}

hr:after {
	content: '&#8711';	
	display: inline-block;
	position: relative;
	top: -15px;
	padding: 0 10px;
	background: #f0f0f0;
	color: #8c8b8b;
	font-size: 18px;
}
</style>

<hr>

To see the R code that generated this webpage, you can see it at my Github page:

<a href="https://github.com/Emmaloto/rstudio-projects" class="w3-btn w3-ripple w3-indigo"> See Code on Github </a>  


<a href="https://icyblaze16.wixsite.com/emmaloto/programming-projects" class="w3-btn w3-ripple w3-teal"> My Website </a> 
